Dissertation

### 2023-10-16 Manuscript/manuscript.qmd
More complete draft of manuscript pdf

---
title: An instance-based model account of the benefits of varied practice in visuomotor skill
date: last-modified
#page-layout: full
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
editor: 
  markdown: 
    wrap: 72
---

[Pdf of the journal
article](Assets/Gorman_Goldstone_2022_Instance-based_model_varied_practice.pdf){target="_blank"}\
[Link to online version of journal
article](https://www.sciencedirect.com/science/article/abs/pii/S0010028522000299){target="_blank"}

```{r}
pacman::p_load(tidyr,papaja, knitr, tinytex, RColorBrewer, kableExtra, cowplot, patchwork)
source('Functions/IGAS_ProcessFunctions.R')

theme_set(theme_classic())
# load the processed data from experiment 1 and 2
e1 <- readRDS("data/igas_e1_cleanedData-final.rds")%>% mutate(initialVelocityX=X_Velocity,initialVelocityY=Y_Velocity,stageInt=as.numeric(as.character(experimentStage)))
e2<- readRDS('data/igas_e2_cleanedData-final.rds')%>% mutate(initialVelocityX=X_Velocity,initialVelocityY=Y_Velocity)
# load subject similarity data - computed with the IGAS model in 'IGAS-SimModel.R'
e2_sim <- readRDS('data/IGAS_Similarity-Performance.rds')

options(contrasts = c("contr.sum", "contr.poly"))
defaultContrasts = options()$contrasts
theme_set(theme_classic())

@@

# Abstract

@@

@@

## Similarity and instance-based approaches to transfer of learning

@@

## Issues with Previous Research

@@

# Experiment 1

### Sample Size Estimation

@@

@@

### Task

The experimental task was programmed in JavaScript, using packages from
the Phaser physics engine (https://phaser.io) and the jsPsych library
(de Leeuw, 2015). The stimuli, presented on a black background,
consisted of a circular blue ball -- controlled by the participant via
the mouse or trackpad cursor; a rectangular green target; a red
rectangular barrier located between the ball and the target; and an
orange square within which the participant could control the ball before
releasing it in a throw towards the target. Because the task was
administered online, the absolute distance between stimuli could vary
depending on the size of the computer monitor being used, but the
relative distance between the stimuli was held constant. Likewise, the
distance between the center of the target, and the training and testing
locations was scaled such that relative distances were preserved
regardless of screen size. For the sake of brevity, subsequent mentions
of this relative distance between stimuli, or the position where the
ball landed in relation to the center of the target, will be referred to
simply as distance. @fig-IGAS_Methods displays the layout of the task, as
it would appear to a participant at the start of a trial, with the ball
appearing in the center of the orange square. Using a mouse or trackpad,
participants click down on the ball to take control of the ball,
connecting the movement of the ball to the movement of the cursor.
Participants can then "wind up" the ball by dragging it (within the
confines of the orange square) and then launch the ball by releasing the
cursor. If the ball does not land on the target, participants are
presented with feedback in red text at the top right of the screen, on
how many units away they were from the center of the target. If the ball
was thrown outside of the boundary of the screen participants are given
feedback as to how far away from the target center the ball would have
been if it had continued its trajectory. If the ball strikes the barrier
(from the side or by landing on top), feedback is presented telling
participants to avoid hitting the barrier. If participants drag the ball
outside of the orange square before releasing it, the trial terminates,
and they are reminded to release the ball within the orange square. If
the ball lands on the target, feedback is presented in green text,
confirming that the target was hit, and presenting additional feedback
on how many units away the ball was from the exact center of the target.

@@

```{r}
#| label: fig-IGAS_Methods
#| fig.cap: The stimuli of the task consisted of a blue ball, which the participants would launch at the green target, while avoiding the red barrier. On each trial, the ball would appear in the center of the orange square, with the position of the orange square varying between experimental conditions. Participants were constrained to release the ball within the square

mf <- cowplot::ggdraw()+cowplot::draw_image("Assets/methodsFig1.png",hjust=0)+theme(plot.margin = margin(0, 0, 0, 0))
plot_grid(mf,ncol=1)

## Data Processing and Statistical Packages

@@

## Training Phase

@fig-IGAS_Training1 below shows aggregate training performance binned into three
stages representing the beginning, middle, and end of the training
phase. Because the two conditions trained from target distances that
were not equally difficult, it was not possible to directly compare
performance between conditions in the training phase. Our focus for the
training data analysis was instead to establish that participants did
improve their performance over the course of training, and to examine
whether there was any interaction between training stage and condition.
Descriptive statistics for the intermittent testing phase are provided
in the supplementary materials.

@@

```{r}
#| label: fig-IGAS_Training1
#| fig.cap: Training performance for varied and constant participants binned into three stages. Shorter bars indicate better performance (ball landing closer to the center of the target). Error bars indicate standard error of the mean.

@@

## Testing Phase

@@

```{r}
#| label: fig-IGAS_Testing1
#| fig.cap: Testing performance for each of the 4 testing positions, compared between training conditions. Positions 610 and 910 were trained on by the varied group, and novel for the constant group. Position 760 was trained on by the constant group, and novel for the varied group. Position 835 was novel for both groups. Shorter bars are indicative of better performance (the ball landing closer to the center of the target). Error bars indicate standard error of the mean.

@@

```{r}
#| label: tbl-IGAS_Table1
#| tab.cap: Testing performance for varied and constant groups in experiment 1. Mean absolute deviation from the center of the target, with standard deviations in parenthesis.

@@

## Discussion

@@

@@

### Data Processing and Statistical Packages

@@

### Training Phase

@@

@@

\
\
\

@@

@@

# Computational Model

@@

## Fitting model parameters separately by group

@@

# General Discussion

@@

Also of interest is whether the IGAS model can predict the pattern of
results wherein the varied condition outperforms the constant condition
even from the position on which the constant condition trained. Although
our models were fit using all of the Experiment 2 training and testing
data, not just that of the identity comparisons, in @fig-Toy_Model2 we
demonstrate how a simplified version of the IGAS model could in
principle produce such a pattern. In addition to the assumption of
differential generalization between varied and constant conditions, our
simplified model makes explicit an assumption that is incorporated into
the full IGAS model -- namely that even when being tested from a
position identical to that which was trained, there are always some
psychological contextual differences between training and testing
throws, resulting in a non-zero dissimilarity.

```{r}
#| label: fig-Toy_Model2
#| fig.cap: A simple model depicting the necessity of both of two separately fit generalization parameters, c, and a positive distance between training and testing contexts, in order for an instance model to predict a pattern of varied training from stimuli 400 and 800 outperforming constant training from position 600 at a test position of 600. For the top left panel, in which the generalization model assumes a single c value (-.008) for  both varied and constant conditions, and identical contexts across training and testing, the equation which generates the varied condition is - Amount of Generalization =  $e^{(c\\cdot|x-800|)}  +e^{(c\\cdot|x-400|)}$, whereas the constant group generalization is generated from $2\\cdot e^{(c\\cdot|x-600|)}$. For the top right panel, the c constants in the original equations are different for the 2 conditions, with $c=-.002$ for the varied condition, and $c=-.008$ for the constant condition. The bottom two panels are generated from identical equations to those immediately above, except for the addition of extra distance (100 units) to reflect the assumption of some change in context between training and testing conditions. Thus, the generalization model for the varied condition in the bottom-right panel is of the form - Amount of Generalization = $e^{(cvaried\\cdot|x-800|)}  +e^{(cvaried\\cdot|x-400|)}$.

@@

## References

>>>>

---
title: "Dissertation"
short-title: "Variability and Generalization"
date: 10-11-2023
author:
- name: Thomas Gorman
  affiliation: Indiana University
  url: www.tegorman13.github.io
  email: tegorman@iu.edu
  orcid: 0000-0001-5366-5442

abstract: |
  In project 1, we applied model-based techniques to quantify and control for the similarity between training and testing experience, which in turn enabled us to account for the difference between varied and constant training via an extended version of a similarity based generalization model. In project 2, we will go a step further, implementing a full process model capable of both 1) producing novel responses and 2) modeling behavior in both the learning and testing stages of the experiment. Project 2 also places a greater emphasis on extrapolation performance following training - as varied training has often been purported to be particularly beneficial in such situations. 
keywords:
  - Learning Generalization
  - Function Learning
  - Visuomotor learning
  - Training Variability
code-repo: "Access the code, data, and analysis at <https://github.com/tegorman13/Dissertation>"
bibliography: ../Assets/Bib/Dissertation.bib
link-citations: true
keep-md: true
execute: 
  warning: false
  eval: true
  include: false
---

```{r setup}
here::set_here(path='..')
source(here::here("Functions", "packages.R"))

if (is.null(knitr::pandoc_to())) {
  fmt_out <- "interactive"
} else {
  fmt_out <- knitr::pandoc_to()
}

knitr::opts_chunk$set(echo = FALSE, include =TRUE, 
                      warning = FALSE, message = FALSE, eval=TRUE)

knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "100%", collapse = TRUE)
# 
options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
#

```{r}
source(here::here('Functions/IGAS_ProcessFunctions.R'))

e1 <- readRDS(here::here("data/igas_e1_cleanedData-final.rds")) %>% mutate(initialVelocityX=X_Velocity,initialVelocityY=Y_Velocity,stageInt=as.numeric(as.character(experimentStage)))

# load the processed data from experiment 1 and 2

e2<- readRDS(here::here('data/igas_e2_cleanedData-final.rds')) %>% mutate(initialVelocityX=X_Velocity,initialVelocityY=Y_Velocity)

# load subject similarity data - computed with the IGAS model in 'IGAS-SimModel.R'
e2_sim <- readRDS(here::here('data/IGAS_Similarity-Performance.rds'))

options(contrasts = c("contr.sum", "contr.poly"))
defaultContrasts = options()$contrasts

@@

# Project 1

## Abstract

@@

## Introduction

@@

### Similarity and instance-based approaches to transfer of learning

@@

### Issues with Previous Research

@@

## Experiment 1

### Methods

#### Sample Size Estimation

@@

#### Participants

@@

#### Task

The experimental task was programmed in JavaScript, using packages from
the Phaser physics engine (https://phaser.io) and the jsPsych library
(de Leeuw, 2015). The stimuli, presented on a black background,
consisted of a circular blue ball -- controlled by the participant via
the mouse or trackpad cursor; a rectangular green target; a red
rectangular barrier located between the ball and the target; and an
orange square within which the participant could control the ball before
releasing it in a throw towards the target. Because the task was
administered online, the absolute distance between stimuli could vary
depending on the size of the computer monitor being used, but the
relative distance between the stimuli was held constant. Likewise, the
distance between the center of the target, and the training and testing
locations was scaled such that relative distances were preserved
regardless of screen size. For the sake of brevity, subsequent mentions
of this relative distance between stimuli, or the position where the
ball landed in relation to the center of the target, will be referred to
simply as distance. @fig-igasTask displays the layout of the task, as
it would appear to a participant at the start of a trial, with the ball
appearing in the center of the orange square. Using a mouse or trackpad,
participants click down on the ball to take control of the ball,
connecting the movement of the ball to the movement of the cursor.
Participants can then "wind up" the ball by dragging it (within the
confines of the orange square) and then launch the ball by releasing the
cursor. If the ball does not land on the target, participants are
presented with feedback in red text at the top right of the screen, on
how many units away they were from the center of the target. If the ball
was thrown outside of the boundary of the screen participants are given
feedback as to how far away from the target center the ball would have
been if it had continued its trajectory. If the ball strikes the barrier
(from the side or by landing on top), feedback is presented telling
participants to avoid hitting the barrier. If participants drag the ball
outside of the orange square before releasing it, the trial terminates,
and they are reminded to release the ball within the orange square. If
the ball lands on the target, feedback is presented in green text,
confirming that the target was hit, and presenting additional feedback
on how many units away the ball was from the exact center of the target.

@@

![The stimuli of the task consisted of a blue ball, which the participants would launch at the green target, while avoiding the red barrier. On each trial, the ball would appear in the center of the orange square, with the position of the orange square varying between experimental conditions. Participants were constrained to release the ball within the square](../Assets/methodsFig1.png){#fig-igasTask}

### Results

#### Data Processing and Statistical Packages

@@

#### Training Phase

@fig-IGAS-Training1 below shows aggregate training performance binned into three
stages representing the beginning, middle, and end of the training
phase. Because the two conditions trained from target distances that
were not equally difficult, it was not possible to directly compare
performance between conditions in the training phase. Our focus for the
training data analysis was instead to establish that participants did
improve their performance over the course of training, and to examine
whether there was any interaction between training stage and condition.
Descriptive statistics for the intermittent testing phase are provided
in the supplementary materials.

@@

```{r}
#| label: fig-IGAS-Training1
#| fig.cap: Training performance for varied and constant participants binned into three stages. Shorter bars indicate better performance (ball landing closer to the center of the target). Error bars indicate standard error of the mean.

@@

@@

```{r}
#| label: fig-IGAS-Testing1
#| fig.cap: Testing performance for each of the 4 testing positions, compared between training conditions. Positions 610 and 910 were trained on by the varied group, and novel for the constant group. Position 760 was trained on by the constant group, and novel for the varied group. Position 835 was novel for both groups. Shorter bars are indicative of better performance (the ball landing closer to the center of the target). Error bars indicate standard error of the mean.

@@

```{r}
#| label: tbl-IGAS-Table1
#| tab.cap: Testing performance for varied and constant groups in experiment 1. Mean absolute deviation from the center of the target, with standard deviations in parenthesis.

@@

### Discussion

@@

## Experiment 2

@@

#### Data Processing and Statistical Packages

@@

#### Training Phase

@@

#### Testing Phase

@@

\

@@

\

@@

### Computational Model

@@

#### Fitting model parameters separately by group

@@

### General Discussion

@@

Also of interest is whether the IGAS model can predict the pattern of
results wherein the varied condition outperforms the constant condition
even from the position on which the constant condition trained. Although
our models were fit using all of the Experiment 2 training and testing
data, not just that of the identity comparisons, in @fig-Toy-Model2 we
demonstrate how a simplified version of the IGAS model could in
principle produce such a pattern. In addition to the assumption of
differential generalization between varied and constant conditions, our
simplified model makes explicit an assumption that is incorporated into
the full IGAS model -- namely that even when being tested from a
position identical to that which was trained, there are always some
psychological contextual differences between training and testing
throws, resulting in a non-zero dissimilarity.

```{r}
#| label: fig-Toy-Model2
#| fig.cap: A simple model depicting the necessity of both of two separately fit generalization parameters, c, and a positive distance between training and testing contexts, in order for an instance model to predict a pattern of varied training from stimuli 400 and 800 outperforming constant training from position 600 at a test position of 600. For the top left panel, in which the generalization model assumes a single c value (-.008) for  both varied and constant conditions, and identical contexts across training and testing, the equation which generates the varied condition is - Amount of Generalization =  $e^{(c\\cdot|x-800|)}  +e^{(c\\cdot|x-400|)}$, whereas the constant group generalization is generated from $2\\cdot e^{(c\\cdot|x-600|)}$. For the top right panel, the c constants in the original equations are different for the 2 conditions, with $c=-.002$ for the varied condition, and $c=-.008$ for the constant condition. The bottom two panels are generated from identical equations to those immediately above, except for the addition of extra distance (100 units) to reflect the assumption of some change in context between training and testing conditions. Thus, the generalization model for the varied condition in the bottom-right panel is of the form - Amount of Generalization = $e^{(cvaried\\cdot|x-800|)}  +e^{(cvaried\\cdot|x-400|)}$.
#| fig.width: 11
#| fig.height: 10

@@

# Project 2

```{r}
test <- readRDS(here::here("data/e1_08-21-23.rds")) |> filter(expMode2 == "Test") 
#options(brms.backend="cmdstanr",mc.cores=4)
e1Sbjs <- test |> group_by(id,condit) |> summarise(n=n())
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)
```

In project 1, we applied model-based techniques to quantify and control for the similarity between training and testing experience, which in turn enabled us to account for the difference between varied and constant training via an extended version of a similarity based generalization model. In project 2, we will go a step further, implementing a full process model capable of both 1) producing novel responses and 2) modeling behavior in both the learning and testing stages of the experiment. Project 2 also places a greater emphasis on extrapolation performance following training - as varied training has often been purported to be particularly beneficial in such situations. Extrapolation has long been a focus of the literature on function learning [@brehmerHypothesesRelationsScaled1974; @carrollFunctionalLearningLearning1963]. Central questions of the function learning literature have included the relative difficulties of learning various functional forms (e.g. linear vs.bilinear vs. quadratic), and the relative effectiveness of rule-based vs. association-based exemplar models vs. various hybrid models [@bottNonmonotonicExtrapolationFunction2004; @deloshExtrapolationSineQua1997; @jonesActiveFunctionLearning2018; @kalishPopulationLinearExperts2004; @mcdanielConceptualBasisFunction2005; @mcdanielPredictingTransferPerformance2009]. However the issue of training variation has received surprisingly little attention in this area.

# Methods

## Participants

Data was collected from 647 participants (after exclusions). The results shown below consider data from subjects in our initial experiment, which consisted of 196 participants (106 constant, 90 varied). The follow-up experiments entailed minor manipulations: 1) reversing the velocity bands that were trained on vs. novel during testing; 2) providing ordinal rather than numerical feedback during training (e.g. correct, too low, too high). The data from these subsequent experiments are largely consistently with our initial results shown below.

## Task

We developed a novel visuomotor extrapolation task, termed the Hit The Wall task, wherein participants learned to launch a projectile such that it hit a rectangle at the far end of the screen with an appropriate amount of force. Although the projectile had both x and y velocity components, only the x-dimension was relevant for the task.  [Link to task demo](https://pcl.sitehost.iu.edu/tg/HTW/HTW_Index.html?sonaid=){target="_blank"}

## Procedure
Upon arrival at the laboratory, participants were provided with a description of the experiment and signed informed consent forms. They were then seated in front of a computer equipped with a mouse and were given instructions on how to perform the "Hit The Wall" (HTW) visuomotor extrapolation task.

The HTW task involved launching projectiles to hit a target displayed on the computer screen. Participants completed a total of 90 trials during the training stage. In the varied training condition, participants encountered three velocity bands (800-1000, 1000-1200, and 1200-1400). In contrast, participants in the constant training condition encountered only one velocity band (800-1000).

During the training stage, participants in both conditions also completed "no feedback" trials, where they received no information about their performance. These trials were randomly interleaved with the regular training trials.

Following the training stage, participants proceeded to the testing stage, which consisted of three phases. In the first phase, participants completed "no-feedback" testing from three novel extrapolation bands (100-300, 350-550, and 600-800), with each band consisting of 15 trials.

In the second phase of testing, participants completed "no-feedback" testing from the three velocity bands used during the training stage (800-1000, 1000-1200, and 1200-1400). In the constant training condition, two of these bands were novel, while in the varied training condition, all three bands were encountered during training.

The third and final phase of testing involved "feedback" testing for each of the three extrapolation bands (100-300, 350-550, and 600-800), with each band consisting of 10 trials. Participants received feedback on their performance during this phase.

Throughout the experiment, participants' performance was measured by calculating the distance between the produced x-velocity of the projectiles and the closest edge of the current velocity band. Lower distances indicated better performance.

After completing the experiment, participants were debriefed and provided with an opportunity to ask questions about the study.

```{dot}
//| label: fig-design-e1
//| fig-cap: "Experiment 1 Design. Constant and Varied participants complete different training conditions."
//| fig-width: 6.0
//| fig-height: 2.5
//| fig-responsive: false
//| column: screen-inset-right
digraph {
  graph [layout = dot, rankdir = LR]

  // define the global styles of the nodes
  node [shape = rectangle, style = filled]

  data1 [label = " Varied Training \n800-1000\n1000-1200\n1200-1400", fillcolor = "#FF0000"]
  data2 [label = " Constant Training \n800-1000", fillcolor = "#00A08A"]
  Test3 [label = "    Final Test \n  Novel With Feedback  \n100-300\n350-550\n600-800", fillcolor = "#ECCBAE"]

  // edge definitions with the node IDs
  data1 -> Test1
  data2 -> Test1

  subgraph cluster {
    label = "Test Phase \n(Counterbalanced Order)"
    Test1 [label = "Test  \nNovel Bands \n100-300\n350-550\n600-800", fillcolor = "#ECCBAE"]
    Test2 [label = "  Test \n  Varied Training Bands  \n800-1000\n1000-1200\n1200-1400", fillcolor = "#ECCBAE"]
    Test1 -> Test2
  }

  Test2 -> Test3
}

## Analyses Strategy

All data processing and statistical analyses were performed in R version 4.31 @rcoreteamLanguageEnvironmentStatistical2020. To assess differences between groups, we used Bayesian Mixed Effects Regression. Model fitting was performed with the brms package in R @burknerBrmsPackageBayesian2017, and descriptive stats and tables were extracted with the BayestestR package @makowskiBayestestRDescribingEffects2019a. Mixed effects regression enables us to take advantage of partial pooling, simultaneously estimating parameters at the individual and group level. Our use of Bayesian, rather than frequentist methods allows us to directly quantify the uncertainty in our parameter estimates, as well as circumventing convergence issues common to the frequentist analogues of our mixed models. For each model, we report the median values of the posterior distribution, and 95% credible intervals.

Each model was set to run with 4 chains, 5000 iterations per chain, with the first 2500 of which were discarded as warmup chains. Rhat values were generally within an acceptable range, with values \<=1.02 (see appendix for diagnostic plots). We used uninformative priors for the fixed effects of the model (condition and velocity band), and weakly informative Student T distributions for for the random effects.

We compared varied and constant performance across two measures, deviation and discrimination. Deviation was quantified as the absolute deviation from the nearest boundary of the velocity band, or set to 0 if the throw velocity fell anywhere inside the target band. Thus, when the target band was 600-800, throws of 400, 650, and 1100 would result in deviation values of 200, 0, and 300, respectively. Discrimination was measured by fitting a linear model to the testing throws of each subjects, with the lower end of the target velocity band as the predicted variable, and the x velocity produced by the participants as the predictor variable. Participants who reliably discriminated between velocity bands tended to have positive slopes with values \~1, while participants who made throws irrespective of the current target band would have slopes \~0.

```{r}
#| label: tbl-e1-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap:
#|   - "Full datasets"
#|   - "Intersection of samples with all labels available"
#| layout-ncol: 1
#| include: true
#| echo: false

result <- test_summary_table(test, "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))

result$constant |>kbl(booktabs = TRUE,
                        linesep = "\\addlinespace[0.5em]",
                        caption = paste("Summary of Deviation- Constant"))
# |>
#   kable_styling(font_size = ifelse(fmt_out == "latex", 8.5, NA))

result$varied |> kbl(booktabs = TRUE,
                        linesep = "\\addlinespace[0.5em]",
                        caption = paste("Summary of Deviation- Varied"))

### Testing Phase - No feedback.

In the first part of the testing phase, participants are tested from each of the velocity bands, and receive no feedback after each throw.

#### Deviation From Target Band

Descriptive summaries testing deviation data are provided in @tbl-e1-test-nf-deviation and @fig-e1-test-dev. To model differences in accuracy between groups, we used Bayesian mixed effects regression models to the trial level data from the testing phase. The primary model predicted the absolute deviation from the target velocity band (dist) as a function of training condition (condit), target velocity band (band), and their interaction, with random intercepts and slopes for each participant (id).

```{=tex}
\begin{equation}
dist_{ij} = \beta_0 + \beta_1 \cdot condit_{ij} + \beta_2 \cdot band_{ij} + \beta_3 \cdot condit_{ij} \cdot band_{ij} + b_{0i} + b_{1i} \cdot band_{ij} + \epsilon_{ij}
\end{equation}
```

```{r}
#| label: fig-e1-test-dev
#| fig-cap: E1. Deviations from target band during testing without feedback stage. 
#| include: true
test |>  ggplot(aes(x = vb, y = dist,fill=condit)) +
    stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
  labs(x="Band", y="Deviation From Target")
```

```{r}
#| label: tbl-e1-bmm-dist
#| tbl-cap: "Experiment 1. Bayesian Mixed Model predicting absolute deviation as a function of condition (Constant vs. Varied) and Velocity Band"
#| tbl-subcap: ["Constant Testing1 - Deviation", "Varied Testing - Deviation"]
#| include: true
#| echo: false
#| eval: true
modelName <- "e1_testDistBand_RF_5K"
e1_distBMM <- brm(dist ~ condit * bandInt + (1 + bandInt|id),
                      data=test,file=paste0(here::here("data/model_cache",modelName)),
                      iter=5000,chains=4)
GetModelStats(e1_distBMM) |> kbl(booktabs = TRUE,caption = paste("Coefficients"))

e1_distBMM |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600,800,1000,1200)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> gather_emmeans_draws()  |> 
   summarize(median_qi(.value),pd=sum(.value>0)/n()) |>
   select(contrast,Band=bandInt,value=y,lower=ymin,upper=ymax,pd) |> 
   mutate(across(where(is.numeric), \(x) round(x, 2)),
          pd=ifelse(value<0,1-pd,pd)) |> kbl(booktabs = TRUE)
# |> 
#   kbl(caption="Contrasts")

coef_details <- get_coef_details(e1_distBMM, "conditVaried")

The model predicting absolute deviation (dist) showed clear effects of both training condition and target velocity band (Table X). Overall, the varied training group showed a larger deviation relative to the constant training group (β = `r coef_details$estimate`, 95% CI \[`r coef_details$conf.low`, `r coef_details$conf.high`\]). Deviation also depended on target velocity band, with lower bands showing less deviation. See @tbl-e1-bmm-dist for full model output.

#### Discrimination between bands

In addition to accuracy/deviation, we also assessed the ability of participants to reliably discriminate between the velocity bands (i.e. responding differently when prompted for band 600-800 than when prompted for band 150-350). @tbl-e1-test-nf-vx shows descriptive statistics of this measure, and Figure 1 visualizes the full distributions of throws for each combination of condition and velocity band. To quantify discrimination, we again fit Bayesian Mixed Models as above, but this time the dependent variable was the raw x velocity generated by participants on each testing trial.

```{=tex}
\begin{equation}
vx_{ij} = \beta_0 + \beta_1 \cdot condit_{ij} + \beta_2 \cdot bandInt_{ij} + \beta_3 \cdot condit_{ij} \cdot bandInt_{ij} + b_{0i} + b_{1i} \cdot bandInt_{ij} + \epsilon_{ij}
\end{equation}
```

```{r}
#| label: fig-e1-test-vx
#| fig-cap: E1 testing x velocities. Translucent bands with dash lines indicate the correct range for each velocity band. 
#| fig-width: 11
#| fig-height: 9
#| include: true
test %>% group_by(id,vb,condit) |> plot_distByCondit()

```{r}
#| label: tbl-e1-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| layout-ncol: 1
#| include: true

result <- test_summary_table(test, "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kbl(booktabs = TRUE)
result$varied |> kbl(booktabs = TRUE)

```{r}
#| label: tbl-e1-bmm-vx
#| tbl-cap: "Experiment 1. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"
#| tbl-subcap: ["Model fit to all 6 bands", "Model fit to 3 extrapolation bands"]
#| include: true
#| eval: true
e1_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
GetModelStats(e1_vxBMM ) |> kbl(booktabs=T, caption="Fit to all 6 bands")

cd1 <- get_coef_details(e1_vxBMM, "conditVaried")
sc1 <- get_coef_details(e1_vxBMM, "bandInt")
intCoef1 <- get_coef_details(e1_vxBMM, "conditVaried:bandInt")

modelName <- "e1_extrap_testVxBand"
e1_extrap_VxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                  data=test |>
                    filter(expMode=="test-Nf"),file=paste0(here::here("data/model_cache",modelName)),
                  iter=5000,chains=4)
GetModelStats(e1_extrap_VxBMM ) |> kable(booktabs=T, caption="Fit to 3 extrapolation bands")

sc2 <- get_coef_details(e1_extrap_VxBMM, "bandInt")
intCoef2 <- get_coef_details(e1_extrap_VxBMM, "conditVaried:bandInt")

See @tbl-e1-bmm-vx for the full model results. The estimated coefficient for training condition ($B$ = `r cd1$estimate`, 95% CrI \[`r cd1$conf.low`, `r cd1$conf.high`\]) suggests that the varied group tends to produce harder throws than the constant group, but is not in and of itself useful for assessing discrimination. Most relevant to the issue of discrimination is the slope on Velocity Band ($B$ = `r sc1$estimate`, 95% CrI \[`r sc1$conf.low`, `r sc1$conf.high`\]). Although the median slope does fall underneath the ideal of value of 1, the fact that the 95% credible interval does not contain 0 provides strong evidence that participants exhibited some discrimination between bands. The estimate for the interaction between slope and condition ($B$ = `r intCoef1$estimate`, 95% CrI \[`r intCoef1$conf.low`, `r intCoef1$conf.high`\]), suggests that the discrimination was somewhat modulated by training condition, with the varied participants showing less senitivity between vands than the constant condition. This difference is depicted visually in @fig-e1-bmm-vx.@tbl-e1-slope-quartile shows the average slope coefficients for varied and constant participants separately for each quartile. The constant participant participants appear to have larger slopes across quartiles, but the difference between conditions may be less pronounced for the top quartiles of subjects who show the strongest discrimination. Figure @fig-e1-bmm-bx2 shows the distributions of slope values for each participant, and the compares the probability density of slope coefficients between training conditions. @fig-e1-indv-slopes

The second model, which focused solely on extrapolation bands, revealed similar patterns. The Velocity Band term ($B$ = `r sc2$estimate`, 95% CrI \[`r sc2$conf.low`, `r sc2$conf.high`\]) still demonstrates a high degree of discrimination ability. However, the posterior distribution for interaction term ($B$ = `r intCoef2$estimate`, 95% CrI \[`r intCoef2$conf.low`, `r intCoef2$conf.high`\] ) does across over 0, suggesting that the evidence for decreased discrimination ability for the varied participants is not as strong when considering only the three extrapolation bands.

```{r}
#| label: fig-e1-bmm-vx
#| fig-cap: Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands. 
#| fig-subcap: ["Model fit to all 6 bands", "Model fit to only 3 extrapolation bands"]
#| layout-ncol: 2
#| include: true

e1_vxBMM |> emmeans( ~condit + bandInt, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |> 
  condEffects() +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(test$vb), 
                     limits = c(0, 1400))

e1_extrap_VxBMM |> emmeans( ~condit + bandInt, 
                       at = list(bandInt = c(100, 350, 600))) |>
  gather_emmeans_draws() |>
  condEffects() +
  scale_x_continuous(breaks = c(100, 350, 600), 
                     labels = levels(test$vb)[1:3], 
                     limits = c(0, 1000))

```{r}
#| label: tbl-e1-slope-quartile
#| tbl-cap: "Slope coefficients by quartile, per condition"

new_data_grid=map_dfr(1, ~data.frame(unique(test[,c("id","condit","bandInt")]))) |> 
  dplyr::arrange(id,bandInt) |> 
  mutate(condit_dummy = ifelse(condit == "Varied", 1, 0))

indv_coefs <- as_tibble(coef(e1_vxBMM)$id, rownames="id")|> 
  select(id, starts_with("Est")) |>
  left_join(e1Sbjs, by=join_by(id) )

fixed_effects <- e1_vxBMM |> 
  spread_draws(`^b_.*`,regex=TRUE) |> arrange(.chain,.draw,.iteration)

random_effects <- e1_vxBMM |> 
  gather_draws(`^r_id.*$`, regex = TRUE, ndraws = 1500) |> 
  separate(.variable, into = c("effect", "id", "term"), sep = "\\[|,|\\]") |> 
  mutate(id = factor(id,levels=levels(test$id))) |> 
  pivot_wider(names_from = term, values_from = .value) |> arrange(id,.chain,.draw,.iteration)

 indvDraws <- left_join(random_effects, fixed_effects, by = join_by(".chain", ".iteration", ".draw")) |> 
  rename(bandInt_RF = bandInt,RF_Intercept=Intercept) |>
  right_join(new_data_grid, by = join_by("id")) |> 
  mutate(
    Slope = bandInt_RF+b_bandInt,
    Intercept= RF_Intercept + b_Intercept,
    estimate = (b_Intercept + RF_Intercept) + (bandInt*(b_bandInt+bandInt_RF)) + (bandInt * condit_dummy) * `b_conditVaried:bandInt`,
    SlopeInt = Slope + (`b_conditVaried:bandInt`*condit_dummy)
  )

  indvSlopes <- indvDraws |> group_by(id) |> median_qi(Slope,SlopeInt, Intercept,b_Intercept,b_bandInt) |>
  left_join(e1Sbjs, by=join_by(id)) |> group_by(condit) |>
    select(id,condit,Intercept,b_Intercept,starts_with("Slope"),b_bandInt, n) |>
  mutate(rankSlope=rank(Slope)) |> arrange(rankSlope)   |> ungroup()

  indvSlopes |> mutate(Condition=condit) |>  group_by(Condition) |> 
    reframe(enframe(quantile(SlopeInt, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "SlopeInt")) |> 
  pivot_wider(names_from=quantile,values_from=SlopeInt,names_prefix="Q_") |>
  group_by(Condition) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kbl()

```{r}
#| label: fig-e1-bmm-bx2
#| fig-cap: Slope distributions between condition
#| fig-subcap: ["Slope estimates by participant - ordered from lowest to highest within each condition. ", "Destiny of slope coefficients by training group"]
#| layout-ncol: 2
#| fig-height: 9
#| fig-width: 10
#| include: true

  indvSlopes |> ggplot(aes(y=rankSlope, x=SlopeInt,fill=condit,color=condit)) + 
  geom_pointrange(aes(xmin=SlopeInt.lower , xmax=SlopeInt.upper)) + 
  labs(x="Estimated Slope", y="Participant")  + facet_wrap(~condit)

   ggplot(indvSlopes, aes(x = SlopeInt, color = condit)) + 
  geom_density() + labs(x="Slope Coefficient",y="Density")

```{r}
#| label: fig-e1-indv-slopes
#| fig-cap: Subset of Varied and Constant Participants with the smallest and largest estimated slope values. Red lines represent the best fitting line for each participant, gray lines are 200 random samples from the posterior distribution. Colored points and intervals at each band represent the empirical median and 95% HDI. 
#| fig-subcap: ["subset with largest slopes", "subset with smallest slopes"]
#| fig-height: 10
#| fig-width: 12
#| include: true
